\subsection{Java}
%-   Intro                       
%  -   Influence                     
%  -   History                     
%  -   Intent                      

The Java virtual machine starts and loads up classes from a
\texttt{ClassLoader}, an object responsible for loading class definitions into
the system for use. \texttt{ClassLoader}s can be custom--written and added to
the system to work in unison with the default file--based loader, allowing
networked access or on--the--fly class generation\cite{javassist}.

Beyond knowing how to load segments of code in classes, Java's virtual machine
has another capability granted from its knowledge of the object model: garbage
collection. The Java language is fully and transparently garbage collected.
Unreferenced objects, even those who keep cycles of references to each other,
are properly destroyed and their memory reclaimed during the normal execution
of a Java program.

Modern virtual machines are used in high--load, mission--critical
environments. They have become very solid and their performance has been
extensively optimized. Today's VMs watch for often--used sequences of Java
bytecode and compile them into native machine instructions for further
optimization. Also, completely--native compilers exist for Java\cite{GCJ} that
eliminate the need for a virtual machine and its overhead.

\subsubsection{Compilation}

Before a class can be loaded, it has to be compiled. Compilation occurs before
program execution. Each class in Java has its own \texttt{.java} file that is
compiled into a \texttt{.class} file. Each \texttt{.java} file can only have
one publicly--accessible class in it, defined with the same name as the
filename. All static checks are performed, inside a class and between classes.
Because classes are compiled into separately--loaded files, additional checks
are needed at runtime to make sure the file is compatible with the system.

In general, run--time compatibility errors show up as accesses to undefined or
inaccessible members or methods of a class. These failed access attempts exist
as exceptions, which can be caught. Due to the need to catch these kinds of
errors, and other reasons, Java saves a rich amount of information about each
compiled class directly within the \texttt{.class} file, which is loaded and
maintained by the virtual machine during execution. Java's introspection
mechanism exposes this information to the programmer.

Even though there isn't a true linking stage within the Java compilation
model, there's still a need to distribute single files for entire applications
or libraries. The Java utility \texttt{jar} acts almost exactly like the
standard Unix \texttt{tar} utility; only that it uses the \texttt{.zip} file
format and stores a manifest file inside the archive. The Java virtual
machine's standard \texttt{ClassLoader} can directly read \texttt{jar}s and
load class code from them.

%  \texttt{jar} files can contain more than just executable code: text files, resources, graphics, and anything else can be saved in them.  The \texttt{ClassLoader} will ignore them, but they can be accessed by the code within the \texttt{jar}.  Such an ability enables the single--file distribution of an entire Java application.  Java's J2EE (Java 2 Enterprise Edition) uses such a distribution model for its web applications, which deploy as a single file that's loaded by the application server and executed.  Java's WebStart application allows a simple URL to specify a remote application, which is loaded on demand and cached locally.  WebStart removes the need for an application to have a traditional installation phase, and always keeps the cached version up--to--date; greatly simplifying desktop maintenance across an organization.

%-   Compilation                   
%  -   Compilation process               
%  -   Preprocessing stages                
%  -   input & output                    
%  -   archives & libraries                  
%  -   application management           

\subsubsection{Syntax}

Java's syntax for basic procedural use is very similar to C and C++. However,
it cannot be used without using the object--oriented features. As implied by
the compilation process, Java only compiles classes. All code must exist
within classes, either in methods, constructors, or static initialization
blocks.

To demonstrate, take a look at Figure \ref{fig:java-simple}. Like C, the entry
point for any program in Java is \texttt{main()}. However, Java doesn't allow
any code to exist outside of a class definition, hence the need for the
otherwise--useless \texttt{MainContainer}.

The syntax for declarations, loops, and conditionals is nearly identical to C
and C++. The standard set of \texttt{for}, \texttt{while}, and \texttt{do}
loops exists. Two main differences exist: (1) statements given as parameters
to conditionals must be of \texttt{boolean} type, and (2) there are no
explicit references or pointers. For example, the comparison \texttt{(i==0)}
could be written as simply \texttt{(i)} in C or C++; if all the bits of
\texttt{i} are zero, then \texttt{i} is considered \texttt{false}. Also, the
variable \texttt{mc} is a reference to a heap--allocated
\texttt{MainContainer}. Unlike C++, no \texttt{*} or \texttt{\&} is necessary.

For primitive data types, the declarations are nearly identical to C and C++.
Every other declaration is implicitly a reference to an object. References to
references are not allowed. Although the references keep the same lifetime as
primitive types, the lifetimes of the objects are different. All objects live
on the heap through a call to \texttt{new}, and die upon garbage collection.

\begin{figure}[ht!]
\begin{verbatim}
public class MainContainer {
  int loop = 32;
  
  public static void main (String args[]) {
    MainContainer mc = new MainContainer ();
  for (int i=0; i<loop; i++) {
    System.out.println ("Hello World");
    if (i == 0) {
      System.out.print ("!-");
    }
  }
  }
}
\end{verbatim}
\caption{Simple Java Program}
\label{fig:java-simple}
\end{figure}

%-   Syntax                      
%  -   Overview                      
%  -   Procedural syntax                 
%  -   Declarations                    
%  -   Conditionals                    
%  -   Loops                     
%  -   Procedure calls                 

\subsubsection{Type System}

Java has taken a pick--and--choose approach to its type system from C++ and
Objective--C. The primitive types come from the latter. Primitive types like
\texttt{int}, \texttt{float}, \texttt{double}, \texttt{boolean}, and
\texttt{char} all exist in Java and behave as they do in C, C++, or
Objective--C. The only difference is that their specific sizes and precisions
are completely defined; unlike Objective--C or C++ where it varies upon the
specific processor architecture.

Similar to Objective--C's \texttt{NSNumber}, Java has fully--fledged class
types that peer the primitives. \texttt{Integer}, \texttt{Float},
\texttt{Double}, \texttt{Boolean}, and \texttt{String} all provide wrappers
around a primitive value, as well as comparison and conversion operations. As
discussed later, Java's containers can only contain full objects, and these
peers allow the primitive types to be used.

Java doesn't have the concept of an explicit pointer or reference type;
instead, every variable declared to be of a class's type is really a reference
to an instance allocated on the heap. Every such variable has to be
initialized with a call to \texttt{new}, and any other variable set to be of
the same value will refer to the same object. While references to objects can
be declared \texttt{final} like primitive types, making them immutable, the
referred objects are always mutable. To simulate a const object as in C++, the
traditional Java method to return immutable objects is to define a subset of
the object's interface with only ``getter" methods that allow the query but
not modification of an object's state. The true object's exposed interface is
a superset of this, and it formally ``implements" it (discussed more later).

Arrays in Java are essentially unidimensional. They hold either primitive
types or references to objects. However, arrays are also objects, and thus
references to arrays can be stored in arrays as well. Through this
double-indirect mechanism, multidimensional arrays are implemented in Java.
The immediate benefit of the flexibility is clear: arrays are simple to
understand and flexible to use. For example, the references to objects could
be polymorphic; allowing further dynamism.

Arrays, like other Java objects, are always mutable and garbage collected.
They also ``know" their length, and can be queried for them as needed. Arrays
also have a peer, \texttt{Array}, which provides similar wrapping facilities
as the other peers.

Java uses a similar type system as C++. All variables are declared to have
some type, either primitive or object. Attempts to assign the variables values
of other, incompatible types are errors. Downcasting from a class to one of
its subclass types is allowed. However, there is little allowance for implicit
conversion; only upcasting. Even narrowing conversions between floating point
types from literals are errors!

\subsubsection{Objects}

Java's object model is a mix between C++ and Objective--C. The methods are
declared and used almost identically to C++ syntax. However, the inheritance
and polymorphic mechanisms within Java function more similarly to
Objective--C.

Messages in Java look and act almost identically to C++: essentially functions
with a hidden \texttt{this} pointer back to the object. An attempt to call a
method not implemented by the message recipient is flagged as a compile--time
error. Furthermore, methods can be overloaded: more than one method can be
declared and defined with the same name, as long as their parameter list
differs.

Unlike C++, default values for parameters, nor the overloading of operators
are allowed. There is one exception: Java's \texttt{String} class has
concatenation operators defined, with definition for all the primitive types
and behavior to call \texttt{toString} on all other objects given as
parameters.

Class types fall into three categories: classes, abstract classes, and
interfaces. Classes have all methods defined, member variables, and are
constructible. They can also inherit from one other class, and
\emph{implement} any number of interfaces. Interfaces only have methods
declared without implementation. Classes implement interfaces by implementing
all of their methods.

Abstract classes are unconstructable objects with one or more methods
unimplemented and marked \texttt{abstract}. Unlike interfaces, they can have
some methods defined for subclass use, but take the role as the only
superclass. Subclasses must implement all abstract methods of their abstract
superclass and any implemented interfaces to be constructable.

For classes that have non-memory resources allocated, a \emph{finalizer} can
be defined, which is run when the object is garbage collected. However, the
specific time of execution, or even a guarantee of execution, isn't provided.

Access control is extremely C++-like: private, public, and package access is
allowed. Access sections like C++ aren't provided; each member and method has
to have its own qualifier listed, otherwise it's assumed to be package--level
access. Inheritance and interface implementation, however, are always public.

Objects can be queried of their type through several ways. First, the
\texttt{instanceof} operator returns a boolean value specifying if the object
is an instance of a specific type or subtype thereof. Next, \texttt{Class}
provides the comparison methods \texttt{isInstance} and
\texttt{isAssignableFrom}, both of which compare compatibility with another
object. Most often, the \texttt{instanceof} operator is used.

As mentioned before, classes can inherit from exactly one superclass. When one
isn't mentioned, it's Java's standard \texttt{Object}. As a consequence, every
object in the Java system inherits from Object. Without multiple inheritance,
questions about diamond inheritance, ambiguous superclass references, and the
like are completely avoided. Also as mentioned before, each compiled class has
a description that's loaded, maintained, and checked by the virtual machine.
Such a description is available to the developer as an instance of type
\texttt{Class}, available through \texttt{Object}'s \texttt{getClass} method.

All methods are implicitly polymorphic; the C++ keyword \texttt{virtual} is
assumed. The Java keyword \texttt{final} will specify a method that cannot be
overridden in base classes. For a method to be overridden, it has to have the
same access level as the original, and the same signature (method name and
parameter types). As upcasting is an implicit conversion, Java objects are
often treated as if they were instances of their base class or an implemented
interface. Method calls to are routed to the closest ancestor's
implementation.

Exceptions are based on C++: a \texttt{try} block containing code that may
throw, one or more \texttt{catch} blocks that handle a specific type of
exception, and specific to Java, a \texttt{finally} block for cleanup code
that runs even if a stack unwind is in progress.

Unlike C++, Java's exceptions only use class types: throwing an integer or
floating point value isn't possible. The virtual machine is a source of many
exceptions as well. Illegal actions, such as trying to call a method on a null
reference or casting an object not of the specified type, are trapped by the
VM and result in exceptions being thrown in the running program. This gives
the running program a reasonable chance to trap the error and continue
execution.

Java has a generics system in its 5.0 beta as of August
2004\cite{java-generics}. Java's compiler has a simple preprocessor that
allows the declaration of generic types. Generic types in Java have a similar
syntax as C++: angle brackets denote parameters to the generic type. The
parameters are used to denote types used for method parameters, member types,
and return types. The compiler will flag attempts to use an instantiation of a
generic type that doesn't match its definition. For example, a generic
container will not allow insertions of objects that aren't instances or
subclasses of its parameter.

Behind the scenes, the generic types have only one instantiation that's shared
between all uses. The parameter type names are all converted to
\texttt{Object}, and compiled as a normal Java class. This way, the
traditional one--to--one mapping of a \texttt{.java} source file and the
compiled \texttt{.class} still exists. Unfortunately, the only gains from the
generics feature are some type safety and reduced need for casting; none of
the more powerful capabilities generics provide in C++ are available in Java.

%-   Type System                   
%  -   Primitive Type Handling (integer, Integer, etc)       
%  -   Arrays                      
%  -   Type Safety/Checking                
%  -   Object Model                    
	%  -   Messaging                   
	%  -   Overloading                   
		%  -   Operators                     
	%  -   Access Control                    
	%  -   Inheritance                   
	%  -   Polymorphism                    
%  -   Exceptions                    
%  -   Generics                      

\subsubsection{Runtime}

As mentioned before, the virtual machine loads Java \texttt{.class} files via
a \texttt{ClassLoader}, which returns a \texttt{Class} object to the VM. At
startup, the VM is given a single class name to load, which must have a
\texttt{public}, \texttt{static} method named \texttt{main} taking a single
parameter: an array of \texttt{String}. That method is run with the command
line options given at the VM's invocation, with VM--specific options removed.

\texttt{main} may spawn any number of threads, which are supported natively by
the VM. Included with the ability to create new threads are in--language
synchronization abilities, such as the ability to make a method
\texttt{synchronized}: callable only from one thread at a time. The virtual
machine provides the threading and enforces the synchronization, even if the
underlying platform doesn't do it natively.

Like C, C++, and Objective--C, the program lives only as long as \texttt{main}
runs, even if other threads are still active when \texttt{main} completes.
During that lifetime, all memory allocated is tracked and managed by the VM.
Objects and graphs thereof with no incoming references are garbage collected
and their memory reclaimed.

One of Java's most powerful assets is the wealth of standard libraries. The
standard library contains nearly 2,000 classes and interfaces in nearly 100
packages. Together, a platform--independent system for developing desktop,
web, and command--line applications exists, with facilities for almost every
common development need.

Due to the easy packaging and distribution of platform--independent code, Java
also has one of the richest 3rd party library communities.

Within the standard libraries lie Java's introspection mechanism. The
mechanism's libraries allow programmatic access to the class information the
VM maintains. As hinted before, the \texttt{Class} type provides the key
interface for accessing a type's information. With it, members, methods,
interface, and superclass information is all available. Moreover,
\texttt{Class} provides a query function for getting the appropriate
\texttt{Class} instance for a type with a specific name. Such an ability,
connected with the \texttt{ClassLoader} mechanism, allows a Java program to
assimilate code that was not available at the original system's compilation,
loading it, linking it, and running it when appropriate.

\texttt{Class} provides a healthy API, we present only the relevant subset for
this discussion. All the \texttt{public} constructors, methods, and members
are accessible via \texttt{getConstructor}, \texttt{getMethod} and
\texttt{getField}. Plural versions of these methods exist that return arrays
of each as well. These methods return \texttt{Constructor}, \texttt{Method},
and \texttt{Field} objects.

\texttt{Constructor} is essentially a factory class. Given the parameters it
needs for initialization, its \texttt{newInstance} will return a newly
constructed instance of the type. It allows querying of the required parameter
types via \texttt{get\-P\-ar\-am\-et\-er-\-T\-yp\-es}, which returns an array
of \texttt{Class} objects. Note that primitive types do have \texttt{Class}
objects, but they must be passed to \texttt{Constructor} as wrapped objects.

\texttt{Method} acts almost as a selector in Objective--C. Similar to
\texttt{Constructor}, it has a \texttt{get\-Par\-am\-et\-er\-T\-yp\-es} method
for accessing the parameters. It also has \texttt{getReturnType} and
\texttt{getName} for getting the full method description. \texttt{invoke} in
\texttt{Method} takes a recipient object and a set of parameters and invokes
the method on the recipient.

\texttt{Field} provides \texttt{get} and \texttt{set} methods for objects and
pairs of these methods for each primitive type. All of them take a recipient
object which contains the member variable in question. The primitive type
pairs are named as \texttt{getInt} and \texttt{setInt} which return and take
primitive types for primitively--typed members.

All three classes, \texttt{Constructor}, \texttt{Method}, and \texttt{Field}
will throw exceptions when they are used inappropriately. Examples include
passing the wrong parameter types, sending to the wrong object, or passing
invalid values. Because the classes must have type--agnostic interfaces, these
errors cannot be caught by the compiler.

%-   Runtime                     
%  -   Startup                     
%  -   Memory Management               
%  -   Libraries                     
%  -   Introspection                   

\subsubsection{Case Study: Serialization}

Java provides built--in serialization\cite{957325,944589,376846}. By
implementing a zero--method interface \texttt{Serializable}, an object can be
serialized. The primitive types can also be serialized.

Use \texttt{Obj\-ect\-Outp\-utStr\-eam}\cite{java-api}, a wrapper around a
normal Java \texttt{Out\-p\-ut\-Str\-eam}, to serialize the object.
\texttt{Obj\-ect\-Outp\-utStr\-eam}'s \texttt{writeObject} method will
serialize the object to the stream, and its peer
\texttt{Obj\-ect\-Inp\-utStr\-eam}'s \texttt{readObject} will deserialize it.

By sitting atop of the standard stream mechanism in Java, serialization works
atop of any byte--stream I/O mechanism. Included in the standard libraries are
files and sockets. The developer may write their own stream classes and send
serialized objects over them with little difficulty.

Serialization is almost completely transparent to the object being marshaled.
The only time a class need worry about serialization is when it's got members
that don't implement \texttt{Serializable}. While most of Java's classes do,
some don't for obvious reasons, like \texttt{Thread}. For these members, the
class must mark them \texttt{transient} in their declaration.

Furthermore, the class may need to know when it's being serialized or
deserialized, so that it can adjust its state. For example, it would have to
reconstruct any \texttt{transient} members that were lost during
serialization. The class can define private \texttt{readObject} and/or
\texttt{writeObject} methods which will be called during the relevant
processes. From there, it can prepare for serialization or fully restore from
it.

Another option in Java is the \texttt{Externalizable} mechanism, which does
less of the work by itself, in exchange for greater control of the
serialization format.


\subsubsection{Case Study: Distributed Objects}

Java provides a basic distributed objects mechanism called Remote Method
Invocation (RMI)\cite{java-rmi}. By implementing a zero--method interface
\texttt{Remote}, an object specifies that it can be remotely invoked.

When a message is sent to a remote object, the parameters are serialized.
Those parameters which implement \texttt{Remote} are given
remotely--accessible identifiers, which are sent in their place. A parameter
that implements neither \texttt{Remote} nor \texttt{Serializable} can't be
sent. From there, the virtual machines interact to transport and dispatch the
message.
