\section{Open Issues}

At the time of this writing there are two open issues with the runtime type
system registration and invalidation. We enumerate these below which require
committee feedback and further investigation for future revisions to the
paper:

\begin{itemize}

 \item \textbf{Dynamic invalidation of statically registered types.} Currently
it is implied that the registration mechanism for statically-linked type
descriptors is implementation-defined, and that the registry mechanism may
vary. Future revisions to the paper should have this addressed based on
feedback from the committee.

 \item \textbf{Type name to descriptor mapping.} As C++ does not define
link-time requirements on type names and how they map to actual objects from
different modules, introducing a mapping based on a type's name may be
problematic. What comes to mind are types with the same name in the anonymous
namespace in two different translation units. Whether there's a name conflict
is unclear, and an implementation may have to disambiguate the type's names
using some other mechanism than just the mangled symbol name.

 \item \textbf{Shared library semantics are platform-dependent and undefined
by the standard.} A lot of the invalidation and shared-library/module specific
information in the paper assumes a consistent behavior of dynamically loaded
libraries. Without clear standards on how shared-libraries work across
platforms much of the guarantees may not be made on all current platforms.
Feedback on the direction for shared/dynamic library semantics and whether
these should be standardized would be greatly appreciated.

\end{itemize}

It is the authors' hope that these issues would be at least discussed if
interest in getting runtime introspection into the standard at some point in
the future.